---
title: "Final Project"
authors: "Tatum Good, Jace Higa, Kendall Leonard"
format: 
  revealjs:
      embed-resources: true
      mainfont: Century Schoolbook
      slide-number: true
editor: visual

---


```{r echo=FALSE}
library(tidyverse)
library(patchwork)  # Add this to enable patchwork

fast <- readRDS(gzcon(url("https://github.com/cd-public/D505/raw/refs/heads/master/dat/craven_train.rds")))
percent_cols <- sapply(fast, function(col) any(grepl("%", col)))
profit <- fast["Revenue 2019 to present"] - fast["BOM Cost"] * fast["Units Sold"]
fast["Profit"] = profit

percent_colnames <- names(fast)[percent_cols]

fast[ , percent_colnames] <- lapply(fast[ , percent_colnames], function(x) {
  as.numeric(gsub("%", "", x))
})

fast <- fast[ , !sapply(fast, function(x) is.factor(x) && length(na.omit(unique(x))) < 2)]

fast_features <- fast %>%
  mutate(
    bel_avg_BOM = `BOM Cost` < quantile(`BOM Cost`, 0.94),
    high_revenue = `Revenue 2019 to present` > quantile(`Revenue 2019 to present`, 0.94),
    log_revenue = log(`Revenue 2019 to present`),
    interaction_cost_units = `BOM Cost` * `Units Sold`,
    high_volume = `Units Sold` > quantile(`Units Sold`, 0.99),
    src_www = src_www < 60,
    direct = `pct_Direct Sales` > 90,
    form = ifelse(`Main Component MFG Process` == "Form", 1, 0),
    finish = ifelse(`Main Component Material` == "Electronics", 1, 0),
    year_min = June > 3
  ) %>%
  select(Profit, bel_avg_BOM, high_revenue, log_revenue, interaction_cost_units, high_volume, src_www, src_external, direct, form, finish, year_min)

fast_features_clean <- na.omit(fast_features)
for (i in 1:ncol(fast_features_clean)) {
  fast_features_clean[[i]] <- ifelse(is.infinite(fast_features_clean[[i]]), 0, fast_features_clean[[i]])
}

```

## What's the point of it all

- Same as everyone else
    - get the smallest RMSE



- But what does that even mean in this context?
  - Answer: have a model who is off by the lowest amount when predicting profit


## Scrub-a-dub-dub (AKA Cleaning the Data)

- A lot of percentages in this data
  - meaning that a lot of numeric data was being seen as textual, making any meaningful feature building next to impossible
  
- So we chopped off the percentages!
  - Yay! Workable data!

## Features We Used {.smaller}

   - bel_avg_BOM
   - high_revenue 
   - log_revenue 
   - interaction_cost_units 
   - high_volume 
   - src_www 
   - direct 
   - form 
   - finish 
   - year_min 

## A detailed look: bel_94_perc_BOM
- Stands for 'Below the 94th Percentile BOM (Bill of Materials)'
- Bill of Materials: A price that consists of all materials needed to assemble a product.
  - The less the materials cost, the more profit is possible.
- Why the 94th percentile?
  - Tested many percentiles but the 94th ended up lowering our RMSE the most.

## A detailed look: high_revenue_products
- Used the 94th percentile as well.
  - Figured that the more potential revenue you make off of a product the more profit is available.
  - Think selling a house vs. selling hotdogs

## A detailed look: log_revenue {.smaller}
- Revenue often has a skewed distribution, where some products generate much higher sales than others. By applying a logarithmic transformation, we make the distribution more symmetric and reduce the impact of extreme values, helping the model generalize better.

```{r echo=FALSE}
# Combine both ggplots side by side using patchwork
p1 <- ggplot(fast, aes(`Revenue 2019 to present`)) +
  geom_histogram()+labs(
    title = "Unmodified Revenue Distribution",
    subtitle = "2019",
    caption = "CravenSpeed",
    x = "Revenue"
    
  )

p2 <- ggplot(fast_features_clean, aes(log_revenue)) +
  geom_histogram()+
  labs(
    title = "Log of Revenue Distribution",
    subtitle = "2019",
    caption = "CravenSpeed",
    x = "Log of Revenue"
    
  )

p1 + p2  # This combines both plots side by side

```



## A detailed look: interaction_cost_units 

## A detailed look: high_volume 
- Used the 99th percentile as well.
  - Think back to the selling a house vs. selling hotdogs analogy
  - If you sell ten thousand hotdogs there is more room for profit which is where this idea came from

## A detailed look: direct 
## A detailed look: form
## A detailed look: finish 
## A detailed look: year_min 
## A detailed look: Design of Model
## A detailed look: RMSE Reveal
## Thank you!
